---
title: "Assessment 1.1"
author: "Meral Erdogan"
date: "09 09 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# RMarkdown ***editing***


[RMarkdown](https://rmarkdown.rstudio.com/lesson-8.html) documents provide quick, reproducible reporting from R. You write your document in markdown and embed executable R code chunks with the knitr syntax. You can update your document at any time by re-knitting the code chunks. You can then convert your document into several common formats.



#PART 1

## Loading Libraries

```{r}
# R libraries to use, load below libraries
library(tidyverse)
library(skimr)
library(knitr)
library(janitor)
library(ggplot2)
library(scales)
library(lubridate)
```




```{r}
#6 Reading data
fastfood<-read.csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2018/2018-09-04/fastfood_calories.csv')
```


```{r}
#7A Displaying first 10 rows using function head
head(fastfood, 10)
```

```{r}
#7B Displaying first 10 rows and 5 variables using function kable
kable(fastfood[1:10, 1:5], caption = "First 10 Rows and first 5 Variables of Fastfood Dataset")
```



```{r}
#8 Showing observations with more than 1000 calories by using function filter
fastfood%>% 
filter(calories>1000) 
```

```{r}
#9 Showing observations with more than 40 in total_fat and more than 80 in total_carb by using function filter
dont_eat_this<-fastfood%>%
filter(total_fat>40 & total_carb>80)
print(dont_eat_this)
```




```{r}
#10 Showing observations for restaurant Mcdonalds by using function filter
fastfood%>%
filter(restaurant == "Mcdonalds")
```



#PART 2

```{r}
#11 Displaying the types of variables by using function skimr
fastfood%>% 
skim()
```



```{r}
#12 Displaying number of observations for each restaurant by using function count
fastfood%>% 
count(restaurant, sort = TRUE) 
```


```{r}
#13 Displaying average number of calories for each restaurant using function groupby and summarise
fastfood%>% 
group_by(restaurant)%>% 
summarise(average_calories=mean(calories))%>%
arrange(desc(average_calories))%>%
ungroup()
```


```{r}
#14 Adding new variable average calories to the data set
fastfood<-fastfood%>% 
group_by(restaurant)%>% 
mutate(average_calories = mean(calories))%>%
ungroup()
```




```{r}
#15 Displaying total fat for each restaurant with data visualization
theme_set(theme_light())
fastfood %>%
mutate(restaurant = fct_reorder(restaurant, total_fat, sum ))%>%
ggplot(aes(x=restaurant,
     y=total_fat, fill=restaurant))+
  geom_col()+
  coord_flip()+
  labs(x = "Restaurant",
       y = "Total fat",
       title = "Total fat for each type of restaurant")
```

Explanation: From this data visualization we can see total fat for each type of fast food restaurant. We can say Burger King has the highest amount of fat followed by Taco Bell. Chick Fil-A has the least amount of fat and looks like it is better option in terms of fat. Data visualization gives us a clear idea of what the information means by giving it visual context through maps or graphs. This makes the data more natural for the human mind to comprehend and therefore makes it easier to identify trends, patterns, and outliers within large data sets.




```{r}
#16 Adding max total fat variable to the data set by using function groupby and mutate
fastfood<-fastfood%>% 
group_by(restaurant)%>% 
mutate(max_total_fat = max(total_fat))%>%
ungroup()
```




```{r}
#17 Data visualization of sugar & protein values in Mcdonalds
fastfood%>%
filter(restaurant == "Mcdonalds")%>%
ggplot(aes(x=sugar,
     y=protein,
     color="red"))+
  geom_point()+
  labs(x = "Sugar",
       y = "Protein",
       title = "Sugar & Protein Values for Mcdonalds")
```
Explanation: In this chart we can see the sugar and protein values for Mcdonalds. Most of the sugar values are in between zero and twenty and similarly most of the protein values are in between zero and fifty. There are some outliers as well. This visualization is crucial in terms of showing the density.



#PART 3

```{r}
#18 Converting variable into a factor
fastfood%>%
mutate(restaurant=as_factor(restaurant))
```
We can convert variables, which we can use them to group by data, into a factor variable. In our case and for now we can convert variable restaurant into a factor variable. 



```{r}
#19 Creating new variable calories type to data set
fastfood<-fastfood%>%
mutate(calories_type=cut_number(calories, n=3, labels=c("low", "med", "high"), ordered_result=TRUE))
```




```{r}
#20 Data Visualization for number of calories type for each restaurant
theme_set(theme_light())
fastfood%>%
  mutate(restaurant = fct_reorder(restaurant, calories, sum))%>%
  ggplot(
 aes(x=calories_type,
     fill=restaurant))+
  geom_bar()+
  labs(x = "Calories Type",
       y = "Count",
       title = "Number of Calories Type for Each Restaurant")+
  facet_wrap(~ restaurant, nrow = 2)
```

Explanation: In this chart we can see the amount of calories type for each restaurant. Taco Bell has the highest amount of calories type counted followed by Subway which is normal these two restaurants have the highest amount of observations. Chick Fil-A has the least amount.



```{r}
#21 Adding percentage of trans_fat in total_fat
fastfood<-fastfood%>%
mutate(trans_fat_percent=trans_fat/total_fat*100)
```


```{r}
#22 Data Visualization for trans_fat in each restaurant
fastfood%>%
ggplot(aes(trans_fat,
     fill=restaurant))+
  geom_dotplot()+
  labs(x = "",
       y = "Trans Fat",
       title = "Distribution of Transfat for Each Restaurant")+
  facet_wrap(~ restaurant, nrow = 2)
```

Explanation: In this chart we can see the distribution of trans fat for each type of restaurant. Sonic is distributed the most. There are some outliers for couple of restaurants. Burger King has an outlier with value 8 which the highest.Subway and Taco Bell have the highest amount of zero trans fat counts. 




```{r}
#23 Calculating average total fat for each restaurant
fastfood%>%
group_by(restaurant)%>%
summarise(average_total_fat=mean(total_fat))%>%
arrange(desc(average_total_fat))%>%
ungroup()
```



```{r}
#24 Data visualization to compare restaurants for total fat
fastfood%>%
  ggplot(
 aes(x=restaurant,
     y= total_fat,
     fill=restaurant)
)+
  geom_boxplot()
```

Explanation: In this chart we can see the distribution of total fat for each type of restaurant. There are some outlier for 6 different restaurants. Overall, we can say they are evenly distributed. Burger King and Sonic have highest values and wider distribution. 


