---
title: "Assignment 1"
author: "Meral Erdogan"
date: "09 09 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# RMarkdown ***editing***


[RMarkdown](https://rmarkdown.rstudio.com/lesson-8.html) documents provide quick, reproducible reporting from R. You write your document in markdown and embed executable R code chunks with the knitr syntax. You can update your document at any time by re-knitting the code chunks. You can then convert your document into several common formats.



#PART 1

# Loading Libraries

```{r}
# R libraries to use, load below libraries
library(tidyverse)
library(skimr)
library(knitr)
library(janitor)
library(ggplot2)
library(scales)
library(lubridate)
```




```{r}
#6 Reading data

fastfood<-read.csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2018/2018-09-04/fastfood_calories.csv')
```


```{r}
#7A Displaying first 10 rows
head(fastfood, 10)
```

```{r}
#7B Displaying first 10 rows and 5 variables using function kable
kable(fastfood[1:10, 1:5])
```



```{r}
#8 Showing observations with more than 1000 calories
fastfood%>% 
filter(calories>1000) 
```

```{r}
#9 Showing observations with more than 40 in total_fat and more than 80 in total_carb 
dont_eat_this<-fastfood%>%
filter(total_fat>40, total_carb>80)
print(dont_eat_this)
```




```{r}
#10 Showing observations for restaurant Mcdonalds
fastfood%>%
filter(restaurant == "Mcdonalds")
```



#PART 2

```{r}
#11 Displaying the types of variables with using function skimr
fastfood%>% 
skim()
```



```{r}
#12 Displaying number of observations for each restaurant
fastfood%>% 
count(restaurant, sort = TRUE) 
```


```{r}
#13 Displaying average number of calories for each restaurant using function groupby 
fastfood%>% 
group_by(restaurant)%>% 
summarise(average_calories=mean(calories))%>%
arrange(desc(average_calories))%>%
ungroup()
```


```{r}
#14 Adding new variable average calories to the data set
fastfood<-fastfood%>% 
group_by(restaurant)%>% 
mutate(average_calories = mean(calories))%>%
ungroup()
```





```{r}
#15 Displaying total fat for each restaurant with data visualization
theme_set(theme_light())
fastfood%>%
mutate(restaurant = fct_reorder(restaurant, total_fat))%>%
ggplot(aes(x=restaurant,
     y=total_fat))+
  geom_col()+
  coord_flip()+
  labs(x = "Restaurant",
       y = "Total fat",
       title = "Total fat for each restaurant")
```

Explanation: From this data visualization we can see the total fat for each fast food restaurant. We can say Burger King has the highest amount of fat followed by Taco Bell. Chick Fil-A has the least amount of fat and looks like it is better option in terms of fat.


```{r}
#16 Adding max total fat variable to the data set
fastfood<-fastfood%>% 
group_by(restaurant)%>% 
mutate(max_total_fat = max(total_fat))%>%
ungroup()
```





```{r}
#17 Data visualization of sugar & protein values in Mcdonalds
fastfood%>%
filter(restaurant == "Mcdonalds")%>%
ggplot(aes(x=sugar,
     y=protein,
     color="red"))+
  geom_point()+
  labs(x = "Sugar",
       y = "Protein",
       title = "Sugar & Protein Values for Mcdonalds")
```
Explanation: In this chart we can see the sugar and protein values of Mcdonalds. Most of the sugar values are in between zero and twenty and similarly most of the protein values are in between zero and fifty. There are some outliers as well. This visualization is crucial in terms of showing the density.



#PART 3

```{r}
#18 Converting variable into a factor
fastfood%>%
mutate(restaurant=as_factor(restaurant))
```

```{r}
#19 Adding new variable calories type to data set
fastfood<-fastfood%>%
mutate(calories_type=cut_number(calories, n=3, labels=c("low", "med", "high"), ordered_result=TRUE))
```


```{r}
#20 Data Visualization for number of calories type for each restaurant
theme_set(theme_light())
fastfood%>%
 mutate(restaurant = fct_reorder(restaurant, calories))%>%
  ggplot(
 aes(x=calories_type,
     fill=restaurant))+
  geom_bar()+
  labs(x = "Calories Type",
       y = "Amount",
       title = "Number of Calories Type for Each Restaurant")+
  facet_wrap(~ restaurant)
```



```{r}
#21 Adding percentage of trans_fat in total_fat
fastfood<-fastfood%>%
mutate(trans_fat_percent=trans_fat/total_fat*100)
```


```{r}
#22 Data Visualization for trans_fat in each restaurant
fastfood%>%
ggplot(aes(x=restaurant,
     y=trans_fat,
     fill=restaurant))+
  geom_boxplot()+
  labs(x = "Restaurant",
       y = "Trans Fat",
       title = "Distribution of Transfat for Each Restaurant")
```


```{r}
#23 Calculating average total fat for each restaurant
fastfood%>%
group_by(restaurant)%>%
summarise(average_total_fat=mean(total_fat))%>%
arrange(desc(average_total_fat))%>%
ungroup()
```



```{r}
#24 Data visualization to compare restaurants for total fat
fastfood%>%
  ggplot(
 aes(x=restaurant,
     y= total_fat,
     fill=restaurant)
)+
  geom_boxplot()
```






