---
title: "Assessment 1.1"
author: "Meral Erdogan - 4634311"
date: "10 08 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# RMarkdown ***editing***



R Markdown documents provide quick, reproducible reporting from R. You write your document in markdown and embed executable R code chunks with the knitr syntax. You can update your document at any time by re-knitting the code chunks. You can then convert your document into several common formats.



[RMarkdown](https://rmarkdown.rstudio.com/lesson-8.html)



```{r}
library(knitr)


#PART 1

#6

fastfood <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2018/2018-09-04/fastfood_calories.csv')

#7
fastfood%>%
head(10)
kable(fastfood[1:10, 1:5])


#8
fastfood%>% 
filter(calories>1000)  


#9
dont_eat_this<-fastfood%>% 
filter(total_fat>40 & total_carb>80)


#10
fastfood%>% 
filter(restaurant=="Mcdonalds")

```





```{r}

#PART 2
library(skimr)

#11
fastfood%>% 
skim()


#12
fastfood%>% 
count(restaurant) 


#13
fastfood%>% 
group_by(restaurant)%>% 
summarise(average_calories=mean(calories))


#14
fastfood%>% 
group_by(restaurant)%>% 
mutate(average_calories = mean(calories)) 


#16
fastfood%>% 
group_by(restaurant)%>% 
mutate(max_total_fat = max(total_fat)) 



```


```{r}

#15
library(tidyverse)
library(scales)
library(lubridate)

theme_set(theme_light())


data(fastfood)
fastfood%>%ggplot(
 aes(x=restaurant,
     y=total_fat,
     color=restaurant,
     size=15
     )
)+
  geom_line()


```
Data visualization gives us a clear idea of what the information means by giving it visual context through maps or graphs. This makes the data more natural for the human mind to comprehend and therefore makes it easier to identify trends, patterns, and outliers within large data sets



```{r}
#17

fastfood_Mcdonalds<-fastfood%>%
filter(restaurant=="Mcdonalds")


data(fastfood_Mcdonalds)
fastfood_Mcdonalds%>%ggplot(
 aes(x=sugar,
     y=protein,
     color="red")
)+
  geom_point()


```



```{r}
#PART 3


#19

fastfood<-fastfood%>%
mutate(calories_type=cut_number(calories, n=3, labels=c("low", "med", "high")))%>%
arrange(calories)



#18
fastfood%>%
mutate(restaurant=as_factor(restaurant))%>%
mutate(calories_type=as_factor(calories_type))




#20

theme_set(theme_light())

fastfood%>%
  mutate(restaurant = fct_reorder(restaurant, calories))%>%
  ggplot(
 aes(x=calories_type,
     fill=restaurant)
)+
  geom_bar()+
  facet_wrap(~ restaurant)




#21
fastfood<-fastfood%>%
mutate(trans_fat_percent=trans_fat/total_fat*100)


#22

fastfood%>%
 mutate(restaurant = fct_reorder(restaurant, trans_fat))%>%
  ggplot(
 aes(x=restaurant,
     y=trans_fat,
     fill=restaurant)
)+
  geom_boxplot()
 


#23

fastfood%>%
group_by(restaurant)%>%
mutate(average_fat=mean(total_fat))



#24
  
fastfood%>%
  ggplot(
 aes(x=restaurant,
     y= total_fat,
     fill=restaurant)
)+
  geom_boxplot()



```





